Computing properties of independent sets by generic programming tensor networks \thanks{\funding{...}}

### 2021-09-16 paper/paper.tex
update appendix D

\begin{figure} 
    \centering
    \includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm}, clip]{figures/fig1.pdf}
    \caption{Benchmark results for computing different properties of independent sets on a random three regular graph with different tensor element types. \red{for each graph size, is there only one graph?}\blue{yes} (a) treewidth versus the number of vertices for the benchmarked graphs. 
    Legends are the data types that we used in the computation, one can find the corresponding computational task in \Tbl{tbl:generictypes}.
    (b) The computing time for calculating the independence number and for counting of the number of independent sets, the number of MISs, and the number of independent sets having size $\alpha(G)$ and $\alpha(G)-1$.
    (c) The computing time for calculating the independence polynomials with different approaches.
    (d) The computing time for enumeration problems, including the enumeration of all independent set configurations, a single MIS configuration, all MIS configurations, all independent set configurations having size $\alpha(G)$ and $\alpha(G)-1$,  with or without bounding the enumeration space.
    \red{suggestions for the figure: 1. the legend seems to overlap the lines, especially the large subfigure. Not sure what's a better way to include the legend there. 2. I see you guys don't typically use markers. We typically use markers to distinguish the different lines, as well as using the color. Especially, for some journals, there is a physical non-color copy, so some requires the figures to be understandable without colors.}
    \blue{If we use markers, the data poitns are too dense. I prefer not adding markers, not much people read papers with kindle or on paper.}
    \blue{I used a smaller fontsize, there are still some overlaps. Does it look better (or worse) now?}
    }
    \label{fig:benchmark}
\end{figure}

@@

\section{Technical guide}\label{sec:technical}
\begin{description}
	\item[\href{https://github.com/under-Peter/OMEinsum.jl}{OMEinsum} and \href{https://github.com/Happy-Diode/OMEinsumContractionOrders.jl}{OMEinsumContractionOrders}] are packages providing the support for Einstein's (or tensor network) notation and contraction order optimizations.
    \texttt{OMEinsumContractionOrders} implements state of art algorithms for finding the optimal contraction order for the an tensor network, including the KaHypar+Greedy~\cite{Gray2021, Pan2021} and local transformation based approaches~\cite{Kalachev2021},
	\item[\href{https://github.com/TensorBFS/TropicalNumbers.jl}{TropicalNumbers} and \href{https://github.com/TensorBFS/TropicalGEMM.jl}{TropicalGEMM}] are packages providing tropical number and efficient tropical matrix multiplication,
	\item[\href{https://github.com/JuliaGraphs/LightGraphs.jl}{LightGraphs}] is a package providing graph utilities, like random regular graph generator,
	\item[\href{https://github.com/JuliaMath/Polynomials.jl}{Polynomials}] is a package providing polynomial algebra and polynomial fitting,
	\item[\href{https://github.com/scheinerman/Mods.jl}{Mods} and \href{https://github.com/JuliaMath/Primes.jl}{Primes}] are packages providing finite field algebra and prime number generations.
\end{description}

@@

\subsection{Set packing}
Set packing is the hypergraph version of the maximum independent set problem, where each set corresponds to a vertex, each element corresponds to a hyperedge.
To solve the set packing problem, we just remove the rank 2 restriction of an edge tensor
\begin{equation}
    B_{v,w,\ldots, z} = \begin{cases}
        1, & v+w+\ldots+z\leq 1,\\
        0, & otherwise
    \end{cases}
\end{equation}

@@

In section~\ref{sec:indpoly}, we show that the independence polynomial can be obtained by solving the linear equation \Eq{eq:lineareq}. Since the coefficients of the independence polynomial can range many orders of magnitude, the round-off errors in fitting can be significant if we use random floating point numbers for $x_{i}$.
In the main text, we propose to use a finite field $\text{GF}(p)$ to circumvent overflow and round-off errors. Here, we give another method based on discrete Fourier transform.
Instead of choosing $x_{i}$ as random numbers, we can choose them such that they form a geometric sequence in the complex domain $x_j = r\omega^j$, where $r \in \mathbb{R}$ and $\omega = e^{-2\pi i/( \alpha(G)+1)}$. The linear equation thus becomes
\begin{equation}
\left(\begin{matrix}
1 & r & r^2 & \ldots & r^{\alpha(G)} \\
1 & r\omega & r^2\omega^2 & \ldots & r^{\alpha(G)} \omega^{\alpha(G)} \\
\vdots & \vdots & \vdots &\ddots & \vdots \\
1 & r\omega^{\alpha(G)} & r^2\omega^{2{\alpha(G)}} & \ldots & r^{\alpha(G)}\omega^{{\alpha(G)}^2}
\end{matrix}\right)
\left(\begin{matrix}
a_0 \\ a_1 \\ \vdots \\ a_{\alpha(G)}
\end{matrix}\right)
= \left(\begin{matrix}
y_0 \\ y_1 \\ \vdots \\ y_{\alpha(G)}
\end{matrix}\right).
\end{equation}

Let us rearrange the coefficients $r^j$ to $a_j$, the matrix on the left side becomes the discrete Fourier transform matrix. Thus, we can obtain the coefficients by inverse Fourier transform $\vec a_r = {\rm FFT^{-1}}(\omega) \cdot \vec y$, where $(\vec a_r)_j = a_j r ^j$.
By choosing different $r$, one can obtain better precision in low independent set size region  ($r<1$) or high independent set size region ($r>1$).

>>>>

\begin{figure} 
    \centering
    \includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm}, clip]{figures/fig1.pdf}
    \caption{Benchmark results for computing different properties of independent sets of a random three regular graph with different tensor element types.
    The time in these plots only includes tensor network contraction process, without taking the just in time compiling time into account.
    \red{for each graph size, is there only one graph?}\blue{yes}
    (a) treewidth versus the number of vertices for the benchmarked graphs. 
    Legends are the data types that we used in the computation, one can find the corresponding computational task in \Tbl{tbl:generictypes}.
    (b) The computing time for calculating the independence number and for counting of the number of independent sets, the number of MISs, and the number of independent sets having size $\alpha(G)$ and $\alpha(G)-1$.
    (c) The computing time for calculating the independence polynomials with different approaches.
    (d) The computing time for enumeration problems, including the enumeration of all independent set configurations, a single MIS configuration, all MIS configurations, all independent set configurations having size $\alpha(G)$ and $\alpha(G)-1$,  with or without bounding the enumeration space.
    \red{suggestions for the figure: 1. the legend seems to overlap the lines, especially the large subfigure. Not sure what's a better way to include the legend there. 2. I see you guys don't typically use markers. We typically use markers to distinguish the different lines, as well as using the color. Especially, for some journals, there is a physical non-color copy, so some requires the figures to be understandable without colors.}
    \blue{If we use markers, the data poitns are too dense. I prefer not adding markers, not much people read papers with kindle or on paper.}
    \blue{I used a smaller fontsize, there are still some overlaps. Does it look better (or worse) now?}
    }
    \label{fig:benchmark}
\end{figure}

@@

\section{Technical guide}\label{sec:technical}
Before showing the codes,
we want to introduce some Julia packages that playing important roles in our code base.

\begin{description}
	\item[\href{https://github.com/under-Peter/OMEinsum.jl}{OMEinsum} and \href{https://github.com/Happy-Diode/OMEinsumContractionOrders.jl}{OMEinsumContractionOrders}] are packages providing the support for Einstein's (or tensor network) notation and contraction order optimizations.
    \texttt{OMEinsumContractionOrders} implements state of art algorithms for finding the optimal contraction order for the an tensor network, including the KaHypar+Greedy~\cite{Gray2021, Pan2021} and local transformation based approaches~\cite{Kalachev2021},
	\item[\href{https://github.com/TensorBFS/TropicalNumbers.jl}{TropicalNumbers} and \href{https://github.com/TensorBFS/TropicalGEMM.jl}{TropicalGEMM}] are packages providing tropical number and efficient tropical matrix multiplication,
	\item[\href{https://github.com/JuliaGraphs/LightGraphs.jl}{LightGraphs}] is a package providing graph utilities, like random regular graph generator,
	\item[\href{https://github.com/JuliaMath/Polynomials.jl}{Polynomials}] is a package providing polynomial algebra and polynomial fitting,
	\item[\href{https://github.com/scheinerman/Mods.jl}{Mods} and \href{https://github.com/JuliaMath/Primes.jl}{Primes}] are packages providing finite field algebra and prime number generations.
\end{description}

@@

\subsection{Set packing}
Set packing is the hypergraph version of the maximum independent set problem, where each set corresponds to a vertex, each element corresponds to a hyperedge.
To solve the set packing problem, we just remove the rank 2 restriction of an edge tensor
\begin{equation}
    B_{v,w,\ldots, z} = \begin{cases}
        1, & v+w+\ldots+z\leq 1,\\
        0, & \text{otherwise}.
    \end{cases}
\end{equation}

@@

In section~\ref{sec:indpoly}, we show that the independence polynomial can be obtained by solving the linear equation \Eq{eq:lineareq}.
Since the coefficients of the independence polynomial can range many orders of magnitude, the round-off errors in fitting can be significant if we use random floating point numbers for $x_{i}$.
In the main text, we propose to use a finite field $\text{GF}(p)$ to circumvent integer overflow and round-off errors.
One drawback of using finite field algebra is its matrix multiplication is less computational efficient compared with floating point matrix multiplication.
Here, we give an alternative method based on discrete Fourier transform with controllable round off errors.
Instead of choosing $x_{i}$ as random numbers, we can choose them such that they form a geometric sequence in the complex domain $x_j = r\omega^j$, where $r \in \mathbb{R}$ and $\omega = e^{-2\pi i/( \alpha(G)+1)}$. The linear equation thus becomes
\begin{equation}
\left(\begin{matrix}
1 & r & r^2 & \ldots & r^{\alpha(G)} \\
1 & r\omega & r^2\omega^2 & \ldots & r^{\alpha(G)} \omega^{\alpha(G)} \\
\vdots & \vdots & \vdots &\ddots & \vdots \\
1 & r\omega^{\alpha(G)} & r^2\omega^{2{\alpha(G)}} & \ldots & r^{\alpha(G)}\omega^{{\alpha(G)}^2}
\end{matrix}\right)
\left(\begin{matrix}
a_0 \\ a_1 \\ \vdots \\ a_{\alpha(G)}
\end{matrix}\right)
= \left(\begin{matrix}
y_0 \\ y_1 \\ \vdots \\ y_{\alpha(G)}
\end{matrix}\right).
\end{equation}

Let us rearrange the coefficients $r^j$ to $a_j$, the matrix on the left side becomes the discrete Fourier transform matrix. Thus, we can obtain the coefficients by inverse Fourier transform $\vec a_r = {\rm FFT^{-1}}(\omega) \cdot \vec y$, where $(\vec a_r)_j = a_j r ^j$.
By choosing different $r$, one can obtain better precision in low independent set size region by choosing $r<1$ or high independent set size region by choosing $r>1$.
